---
title: "Projec 4 - Data 643"
author: "Ann Liu-Ferrara"
date: "June 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Accuracy and Beyond

1. Compare the accuracy of at least two recommender system algorithms against your offline data.
2. Implement support for at least one business or user experience goal such as increased serendipity, novelty, or diversity.
3. Compare and report on any change in accuracy before and after you've made the change in #2.
4. As part of your textual conclusion, discuss one or more additional experiments that could be performed and/or metrics that could be evaluated only if online evaluation was possible. Also, briefly propose how you would design a reasonable online evaluation environment.

```{r}
library(recommenderlab)
library(ggplot2)
data("MovieLense")

# data selection 
(ratings_movies <- MovieLense[rowCounts(MovieLense) > 50, colCounts(MovieLense) > 100])

# split data
precentage_training <- .8

# setting parameters
# check the minimum # of items purchased by user 
# so there will not be any users without items to test the models
min(rowCounts(ratings_movies))
items_to_keep <- 15

# minimum rating that is considered good
rating_threshold <- 3
# # of run the evaluation
n_eval <- 1

eval_sets <- evaluationScheme(
                data = ratings_movies,
                method = 'split',
                train = precentage_training, 
                given = items_to_keep,
                goodRating = rating_threshold,
                k = n_eval
             )

eval_sets

getData(eval_sets, 'train')
(nrow(getData(eval_sets, 'train'))/nrow(ratings_movies))
getData(eval_sets, 'known')
getData(eval_sets, 'unknown')
nrow(getData(eval_sets, 'known'))/nrow(ratings_movies)
unique(rowCounts(getData(eval_sets, 'known')))
qplot(rowCounts(getData(eval_sets, 'unknown'))) + 
  geom_histogram(binwidth = 10) +
  ggtitle('Unknown items by the users')

# k-fold tests on each user
n_fold <- 4
eval_sets <- evaluationScheme(
  data = ratings_movies, 
  method = 'cross-validation',
  k = n_fold,
  given = items_to_keep,
  goodRating = rating_threshold)
size_sets <- sapply(eval_sets@runsTrain, length)
size_sets


```